Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-05-03 inc/introduction.tex
typo fixes in introduction material

Improvements in the performance of DTEMs, especially ultrafast electron microscopes (UEMs), and UED systems will likely require the implementation of techniques that offset or compensate for deleterious space-charge effects~\cite{armstrong_prospects_2007,lobastov_four-dimensional_2005}.
Such performance optimization efforts will require efficient models of electron pulse propagation dynamics that include the ability to simulate the action of electron optical elements (including the photoelectron gun itself) on the pulse.
Indeed $2^{\text{nd}}$ generation UED instruments are now beginning to employ both magnetic lenses and RF cavities~\cite{oudheusden_electron_2007,veisz_hybrid_2007} to optimize the electron pulse delivery to the specimen; that is, compensate for global space-charge effects in both the transverse and longitudinal pulse dimensions, respectively.
Similarly, future UEM column designs are expected to use RF pulse compression cavities while circumventing high charge density beam cross-overs (i.e., beam foci).

In Chapter \ref{chap:ms_model}, I will introduce the of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing free-space electron bunch propagation in a mean-field self-similar Analytic Gaussian (AG) pulse treatment.
For free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
Due primarily to the versatility of the AG model, which results from its use of transverse and longitudinal pulse position and momentum variances, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
Unfortunately, the AG model as laid out by Michalik and Sipe lacks the ability to model the dynamics of electron pulses under influence of external forces; this prevents the AG model from being used to simulate the behavior of electron pulses in realistic microscope columns.
Additionally, to be useful for UEM, for which the initial electron pulse conditions are of notable importance for the performance of the instrument, a realistic set of initial conditions is essential for accurately modeling the pulse dynamics.

In Chapter \ref{chap:extension}, I present some optimizations to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
I then present my extensions to the Michalik and Sipe AG model which permit the inclusion of generic linear external forces, such as those due to perfect electron-optical elements in a microscope column.
In order to adapt this extended AG model for simulating pulses in a realistic UEM column, I then present specific contributions that arise from magnetic lenses, RF cavities and DC accelerators.
The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
These analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.

In Chapter \ref{chap:model_results}, I apply the extended AG model to draw some initial insight into the dynamics of ultrafast electron pulses traversing a UEM column.
One of the conclusions that I will make is that oblate, or ``disk-like'', electron pulses will be generally preferred to prolate ``cigar-like.'' 
Indeed for UEM, with typical laser pulse durations in the femtosecond to picosecond timescales, the electron pulses are likely to be oblate for reasonably sized incident laser spot sizes.
Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion~\cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance~\cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.
First, that the electron pulses should be emitted from a large-area source, and second, that every effort must be made to reduce the initial rms transverse momentum (i.e., the emission cone angle) of the electron pulse as generated from the photocathode.

All of these considerations motivate the construction of the prototype UEM column being built by our group at UIC.
In Chapter \ref{chap:prototype}, I detail the design and fabrication of large-aperture electron-optical elements, which can readily admit these large width electron pulses; included among these are the accelerator and magnetic lenses.
Then, before I conclude, I will in Chapter \ref{chap:photocathode} highlight the photocathode engineering work which has been undertaken in an attempt to reduce the rms transverse momentum of the emitted electron pulses.
In fact, the search for a low transverse momentum electron source has applications beyond UEM and UED; the fields of X-ray Free Electron Lasers (XFELs,~\cite{nemeth_high_2010}) and Compact X-ray Light Sources (CXLS) \cite{cxls_workshop_report_2012} will also benefit from such a source.
I will present investigations of several different photoemission processes, some of which are yielding very promising results.
In fact, our recent work on excited-state thermionic emission~\cite{berger_excited_2012} may even challenge long-standing ideas about which materials and processes may hold the keys to low-emittance (or high-brightness) electron sources.

>>>>

Improvements in the performance of DTEMs, especially ultrafast electron microscopes (UEMs), and UED systems will likely require the implementation of techniques that offset or compensate for deleterious space-charge effects~\cite{armstrong_prospects_2007,lobastov_four-dimensional_2005}.
Such performance optimization efforts will require efficient models of electron pulse propagation dynamics that include the ability to simulate the action of electron optical elements (including the photoelectron gun itself) on the pulse.
Indeed $2^{\text{nd}}$ generation UED instruments are now beginning to employ both magnetic lenses and radio-frequency (RF) resonant cavities~\cite{oudheusden_electron_2007,veisz_hybrid_2007} to optimize the electron pulse delivery to the specimen; that is, compensate for global space-charge effects in both the transverse and longitudinal pulse dimensions, respectively.
Similarly, future UEM column designs are expected to use RF pulse compression cavities while circumventing high charge density beam cross-overs (i.e., beam foci).

In Chapter \ref{chap:ms_model}, I will introduce the of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing free-space electron bunch propagation in a mean-field self-similar Analytic Gaussian (AG) pulse treatment.
For free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
Due primarily to the versatility of the AG model, which results from its use of transverse and longitudinal pulse position and momentum variances, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
Unfortunately, the AG model as laid out by Michalik and Sipe lacks the ability to model the dynamics of electron pulses under the influence of external forces; this prevents the AG model from being used to simulate the behavior of electron pulses in realistic microscope columns.
Additionally, to be useful for UEM, for which the electron pulse's initial conditions are of notable importance for the performance of the instrument, a realistic set of initial conditions is essential for accurately modeling the pulse dynamics.

In Chapter \ref{chap:extension}, I present some optimizations to the AG model and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
I then present extensions to the Michalik and Sipe AG model which permit the inclusion of generic linear external forces, such as those due to perfect electron-optical elements in a microscope column.
In order to adapt this extended AG model for simulating pulses in a realistic UEM column, I then present specific contributions that arise from magnetic lenses, RF cavities, and DC accelerators.
The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
These analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.

In Chapter \ref{chap:model_results}, I apply the extended AG model to draw some initial insight into the dynamics of ultrafast electron pulses traversing a UEM column.
One of the conclusions that I will make is that electron pulses which are oblate or ``disk-like'' in shape will be generally preferred to those that are prolate or ``cigar-like.'' 
Indeed for UEM, with typical laser pulse durations in the femtosecond to picosecond timescales, the electron pulses are likely to be oblate for reasonably sized incident laser spot sizes.
Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion~\cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance~\cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.
First, that the electron pulses should be emitted from a large-area source, and second, that every effort must be made to reduce the initial rms transverse momentum (i.e., the emission cone angle) of the electron pulse as generated from the photocathode.

All of these considerations motivate the construction of the prototype UEM column being built by our group at UIC.
In Chapter \ref{chap:prototype}, I detail the design and fabrication of large-aperture electron-optical elements, which can readily admit these large width electron pulses; included among these are the accelerator and magnetic lenses.
Then, before I conclude, I will in Chapter \ref{chap:photocathode} highlight the photocathode engineering work which has been undertaken in an attempt to reduce the rms transverse momentum of the emitted electron pulses.
In fact, the search for a low transverse momentum electron source has applications beyond UEM and UED; the fields of X-ray Free Electron Lasers (XFELs,~\cite{nemeth_high_2010}) and Compact X-ray Light Sources (CXLS) \cite{cxls_workshop_report_2012} will also benefit from such a source.
I will present investigations of several different photoemission processes, some of which are yielding very promising results.
Our recent work on excited-state thermionic emission~\cite{berger_excited_2012} may even challenge long-standing ideas about which materials and processes may hold the keys to low-emittance (or high-brightness) electron sources.

### 2013-05-03 main.tex
typo fixes in introduction material

Finally, I want to thank the funding agencies which have made this research possible: 
the National Science Foundation (DMR-0619573) and the Department of Energy (DE-FG52-06NA26213 and DE-FG52-09NA29451), as well as a Department of Education GAANN fellowship (DED P200A070409).

@@

\summary
Dynamic Transmission Electron Microscopy, and its picosecond/femtosecond subclass Ultrafast Electron Microscopy, is an emerging field in instrumentation science.
It attempts to combine the nanoscale spatial resolution of transmission electron microscopes with the temporal resolution of modern ultrafast lasers.
In this thesis I present my contributions to this young field. 
These include a novel model for simulating the dynamics of ultrafast electron pulses in electron microscope systems,
design criteria for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron pulse generation.
I also present the prototype ultrafast electron microscope system being built at UIC.

>>>>

Finally, I want to thank the funding agencies which have made this research possible: 
the National Science Foundation (DMR-0619573) and the Department of Energy (DE-FG52-06NA26213 and DE-FG52-09NA29451).
I have been honored to be personally funded by a Department of Education GAANN fellowship (DED P200A070409), a UIC Dean's Scholar Award fellowship, and a Physics Department James Kouvel Fellowship.

@@

\summary
Dynamic Transmission Electron Microscopy, and its picosecond/femtosecond subclass Ultrafast Electron Microscopy, is an emerging field in instrumentation science.
It attempts to combine the nanoscale spatial resolution of transmission electron microscopes with the temporal resolution of modern ultrafast lasers.
In this thesis, I present my contributions to this young field. 
These include a novel model for simulating the dynamics of ultrafast electron pulses in electron microscope systems,
design criteria for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron pulse generation.
I also present the prototype ultrafast electron microscope system being built at UIC.
